% WAI-MW.tex
\begin{hcarentry}{WAI}
\label{wai}
\report{Michael Snoyman}%11/14
\participants{Greg Weber}
\status{stable}
\makeheader

The Web Application Interface (WAI) is an interface between Haskell web
applications and Haskell web servers. By targeting the WAI, a web framework or web application gets access to multiple deployment platforms. Platforms in use include CGI, the Warp web server, and desktop webkit.

WAI is also a platform for re-using code between web applications and web frameworks through WAI middleware and WAI applications. WAI middleware can inspect and transform a request, for example by automatically gzipping a response or logging a request. The Yesod~\cref{yesod} web framework provides the ability to embed arbitrary WAI applications as subsites, making them a part of a larger web application.

By targeting WAI, every web framework can share WAI code instead of wasting
effort re-implementing the same functionality. There are also some new web
frameworks that take a completely different approach to web development that
use WAI, such as webwire (FRP), MFlow (continuation-based) and dingo (GUI).
The Scotty~\cref{scotty} web framework also continues to be developed, and
provides a lighter-weight alternative to Yesod. Other frameworks- whether
existing or newcomers- are welcome to take advantage of the existing WAI
architecture to focus on the more innovative features of web development.

WAI applications can send a response themselves. For example, wai-app-static is used by Yesod to serve static files. However, one does not need to use a web framework, but can simply build a web application using the WAI interface alone. The Hoogle %\cref{hoogle} 
web service targets WAI directly.

Since the last HCAR, WAI has successfully released version 3.0, which removes dependency on any specific streaming data framework. A separate wai-conduit package provides conduit bindings, and such bindings can easily be provided for other streaming data frameworks.

The WAI community continues to grow, with new applications and web frameworks continuing to be added. We've recently started a new mailing list to discuss WAI related topics. Interested parties are strongly encouraged to join in!

\FurtherReading
\begin{compactitem}
\item\url{http://www.yesodweb.com/book/wai}
\item\url{https://groups.google.com/d/forum/haskell-wai}
\end{compactitem}
\end{hcarentry}
